---
title: Geometry
author: nige
date: '2018-11-08'
bibliography: ["packageCitations.bib"]
cover: "geometry.png"
categories:
  - Bike
  - R
  - Data
tags:
  - data
  - bikes
  - geometry
---

```{r setup, include = F}

  packages <- c("tidyverse"
                , "bookdown"
                , "knitr"
                , "ggbiplot"
                , "DT"
                )
  
  purrr::walk(packages, library, character.only = TRUE)
  
  write_bib(packages,file="packageCitations.bib",tweak=TRUE)
  
  knitr::opts_chunk$set(echo = F
                        , warning = F
                        , message = F
                        #, include = F
                        , collapse = TRUE
                        )
  
  options(knitr.kable.NA = ''
          , knitr.table.format = "html"
          )


```

```{r functions}

  noNA <- function(x) sum(is.na(x)) == 0

```

```{r data}

  dat <- read_csv("../../data/BikeGeometry.csv") %>%
    dplyr::filter(Size >53
                  , Size < 56
                  , !is.na(Trail)
                  )

```

---

Baffled by bike geometry? So am I, so don't look for answers here. But here's how I've helped myself scratch the surface of understanding...

---

The main guide for the important numbers to chase was this [CylingTips article](https://cyclingtips.com/2011/02/the-geometry-of-bike-handling/). Gaps in geometry were filled in using this [bike geometry calculator](https://www.bikegeocalc.com/#7Unnameda0b4c743.9802d270e600.64877f769.85608g971.72104h412.76493i340j340k1335l813.03517m1133.98576n679.88726o1177.24813p889.84772q1213.32131r1345.15065s893.89606t31.8u40v6w2x10y280z40A0B150C25D370E45F172.5G30H30Z). There's also [geometrygeeks.com](https://geometrygeeks.bike/understanding-bike-geometry) providing good explanations of each term.

Analysing geometries accumulated from the web gives Figure \@ref(fig:pca). The (blue) roadies cluster together and appear correlated with high head angle numbers and low chain stay numbers. Then there's a smear of endurance/gravel bikes (in brown and green respectively). The cx bikes (in red) are spread on one half of the figure, apparently correlated with high fork rake numbers.

Focussing on just the gravel (and endurance) bikes gives Figure \@ref(fig:pcaGravel)

---

```{r pca, fig.cap = "Principal components analysis of key bike geometry values"}

  datForPCA <- dat %>%
    dplyr::filter(Bars == "Drop"
                  , Year >= 2018
                  )

  pcDat <- princomp(datForPCA %>%
                      dplyr::select_if(noNA) %>%
                      dplyr::select_if(is.numeric) %>%
                      dplyr::select(SeatAngle
                                    ,HeadAngle
                                    ,BBDrop
                                    ,ChainStay
                                    ,ForkRake
                                    ,Stack
                                    ,Reach
                                    #,HeadTube
                                    ,Trail
                      ) %>%
                      scale(center=FALSE,scale=TRUE)
                      )

  ggbiplot(pcDat
           , labels = paste0(datForPCA$Frame,"_",substr(datForPCA$Year,3,4))
           , groups = datForPCA$Use
           , labels.size = 3
           )

```

---

```{r bikeGeometry}

 datatable(datForPCA %>%
                      dplyr::select(Frame
                                    , Year
                                    , Use
                                    , SeatAngle
                                    , HeadAngle
                                    , BBDrop
                                    , ChainStay
                                    , ForkRake
                                    , Stack
                                    , Reach
                                    , Trail
                      )
           )

```

---

```{r pcaGravel, fig.cap = "PCA for just the grave/endurance bikes"}

  datForPCAGravel <- dat %>%
    dplyr::filter(Bars == "Drop"
                  , grepl("Gravel|Endurance",Use)
                  , Year >= 2018
                  )

  pcDatGravel <- princomp(datForPCAGravel %>%
                      dplyr::select_if(noNA) %>%
                      dplyr::select_if(is.numeric) %>%
                      dplyr::select(SeatAngle
                                    ,HeadAngle
                                    ,BBDrop
                                    ,ChainStay
                                    ,ForkRake
                                    ,Stack
                                    ,Reach
                                    #,HeadTube
                                    ,Trail
                      ) %>%
                      scale(center=FALSE,scale=TRUE)
                      )

  ggbiplot(pcDatGravel
           , labels = paste0(datForPCAGravel$Frame,"_",substr(datForPCAGravel$Year,3,4))
           , groups = datForPCAGravel$Use
           , labels.size = 3
           )

```

# Acknowledgements

```{r packages}
 
  cits <- tibble(cits = read_lines("packageCitations.bib"))
 
  kable(cits %>%
          dplyr::filter(grepl("@",cits)
                        , !grepl("CiteR",cits)
                        ) %>%
          dplyr::rename(Package = cits) %>%
          dplyr::bind_cols(cits %>%
                             dplyr::filter(grepl("title",cits)) %>%
                             dplyr::rename(Description = cits)
                           ) %>%
          dplyr::mutate(Package = gsub("@Manual\\{|,|R-","",Package)
                        , Description = gsub("  title = \\{","",Description)
                        , Description = gsub("\\},","",Description)
                        ) %>%
          dplyr::mutate(Citation = paste0("@R-",Package))
   , caption = "R packages explicitly loaded (i.e. not including dependencies) to create this post"
   )
 
```

---

# Citations 