---
title: Gearing
author: nige
date: '2019-05-28'
slug: gearing
categories:
  - Bike
  - Bikeepacking
  - Gravel
tags:
  - bikes
  - gravel
  - gearing
---

# Front chain ring for 1*11-46

Given an 11-46 cluster, what is a good front chain ring for a 1-by setup?

```{r setup, include = F}
  
  packages <- c("tidyverse"
                , "bookdown"
                , "knitr"
                )

  purrr::walk(packages, library, character.only = TRUE)
  
  write_bib(packages,file="packageCitations.bib",tweak=TRUE)
  
  knitr::opts_chunk$set(echo = F
                        , warning = F
                        , message = F
                        #, include = F
                        , collapse = TRUE
                        , out.width = "110%"
                        )
  
  options(knitr.kable.NA = ''
          , knitr.table.format = "html"
          )

```

```{r data}

  upperThreshold <- 45
  lowerThreshold <- 6

  dat <- tibble(front = seq(26,50,2)) %>%
    dplyr::group_by(front) %>%
    tidyr::expand(rear = c(11,13,15,17,19,21,24,28,32,37,46)) %>%
    dplyr::group_by(front,rear) %>%
    tidyr::expand(setup = "700c * 38mm",diaInches =27.5,diaM = .698) %>%
    dplyr::group_by_all() %>%
    tidyr::expand(cadence = seq(50,110,10)) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(roll = diaM*pi*(front/rear)
                  , speed = roll/1000*cadence*60
                  , lower = if_else(speed <= lowerThreshold,1,0)
                  , upper = if_else(speed >= upperThreshold,1,0)
                  , both = map2_dbl(lower,upper,sum)
                  )
    

```

Figure \@ref(fig:plot) shows front chain ring, rear cog, speed and cadence change in relation to an upper and lower speed target. Targets were chosen as `r lowerThreshold` km/hr as close to walking speed and `r upperThreshold` km/hr as it's rare I travel faster than that.

Cadence was chosen in steps of 10 between `r min(dat$cadence)` rpm and `r max(dat$cadence)` rpm.

Any front chain ring between `r dat %>% dplyr::filter(speed >= upperThreshold) %>% dplyr::filter(front == min(front)) %>% dplyr::pull(front)` and `r dat %>% dplyr::filter(speed <= lowerThreshold) %>% dplyr::filter(front == max(front)) %>% dplyr::pull(front)` cross both lines, given the above constraints. For some reason I'm favouring high-end speed over low-end cadence so I think I'll pick 42.

---

```{r plot, fig.height = 10, fig.cap = "Only a few front chain ring sizes cross both speed thresholds when paired with 11-46"}

  ggplot(dat, aes(rear,speed,colour=cadence)) +
    geom_hline(aes(yintercept = lowerThreshold),linetype=2) +
    geom_hline(aes(yintercept = upperThreshold),linetype=2) +
    geom_line(data = dat %>% dplyr::filter(cadence == 80)) +
    geom_point() +
    facet_wrap(~front) +
    scale_colour_viridis_c() +
    labs(title = ""
         , y = "speed km/hr"
         , x = "rear cog"
         ) +
    #scale_y_log10() +
    theme_dark() +
    labs(title = "Change in speed by rear-cog and cadence. Facets are front chain ring teeth"
         , subtitle = "Dotted lines are bottom and top speed targets of 6 km/hr and 50 km/hr"
         )

```

---

# Acknowledgements

R packages explicitly loaded (i.e. not including dependencies) to create this post are listed in Table \@ref(tab:packages).

```{r packages}
 
  cits <- tibble(cits = read_lines("packageCitations.bib"))
 
  kable(cits %>%
          dplyr::filter(grepl("@",cits)
                        , !grepl("CiteR",cits)
                        ) %>%
          dplyr::rename(Package = cits) %>%
          dplyr::bind_cols(cits %>%
                             dplyr::filter(grepl("title",cits)) %>%
                             dplyr::rename(Description = cits)
                           ) %>%
          dplyr::mutate(Package = gsub("@Manual\\{|,|R-","",Package)
                        , Description = gsub("  title = \\{","",Description)
                        , Description = gsub("\\},","",Description)
                        ) %>%
          dplyr::mutate(Citation = paste0("@R-",Package))
   , caption = "R packages loaded"
   )
 
```

---

# References 